Iroko: Predictive Datacenter Congestion Control

### 2017-12-11 Documents/Report/intro.tex

Developments in the past decade have changed the general networking 
environment. Data centers have emerged as an exciting new research frontier, 
posing novel design challenges and opportunities. Driven by minimization of 
costs and maximization of compute power, data centers must run at the highest 
possible utilization to achieve the ideal compute/cost ratio.
Optimizing the distribution of traffic and simultaneously guaranteeing fairness 
is a perennial challenge for any network operator.

Inefficient routing and congestion control can quickly lead to 
bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load 
network, requiring more sophisticated solutions. In fact, recent insights into 
datacenter network traffic have shown that buffer overruns and small 
short-lived bursts are the primary factor of loss.\cite{fb_dc,msr_dc,dctcp}
Despite the innovation potential of data centers, TCP has dominated as 
the congestion control algorithm of choice.
TCP is a \textit{reactive protocol}, responding to indicators of packet drops
latency. However, this surge of loss and latency in the network already 
portends a problem. Overflowing queues in forwarding elements or mismatched 
hardware capabilities imply that traffic has not been optimally managed.
Ideally, a network should always be "zero-queue", i.e., latency 
will merely be induced by propagation, and not queuing delay.

TCP's algorithm as best-effort solution to globally maximize utilization has 
not fundamentally changed over the last thirty years. However, new research 
advancement raise the question if TCP and its assumptions about networks are 
still a viable option.~\cite{pcc,bbr,perc}

@@

In parallel, admission-control-based approaches have gained traction as a 
viable alternative to the conventional burst-and-backoff mechanism of 
TCP.~\cite{expresspass, fastpass, perc}
While idea of admission control and service guarantees in networks is not 
new~\cite{access_limit, access_limit2}, these designs traditionally 
aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
and untrusted environments such as the internet. In a data center, these 
conditions do not apply. End-hosts are generally considered reliable and 
restricted in behaviour, which allows for great simplification of enforcement 
and prioritization policies. Admission control systems can give administrators 
the benefit of tight control over the data center, mitigating the effects of 
queue-buildup and bursts.

Although it is unclear how predictable data center traffic truly is, various 
works have shown that repeated patterns may 
exist.\cite{msr_dc,fb_dc,traffic,microte}
If data center traffic exhibits predictable patterns, it may be 
possible to devise a proactive mechanism which predicts the traffic matrix of 
next iterations, and adjusts traffic accordingly.

>>>>

Developments in the past decade have changed the general networking 
environment. Data centers have emerged as an exciting new research frontier, 
posing novel design challenges and opportunities. Driven by minimization of 
costs and maximization of compute power, data centers must run at the highest 
possible utilization to achieve the ideal compute/cost ratio.
Optimizing the distribution of traffic and simultaneously guaranteeing fairness 
is a perennial challenge for any network operator.
Inefficient routing and congestion control can quickly lead to 
bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load 
network, requiring more sophisticated solutions. In fact, recent insights into 
datacenter network traffic have shown that buffer overruns and small 
short-lived bursts are the primary factor of loss.\cite{fb_dc,msr_dc,dctcp}

Despite the innovation potential of data centers, TCP has dominated as 
the congestion control algorithm of choice.
Still, TCP is a \textit{reactive protocol}, only responding to signals of 
congestion. A substantial and noticeable increase of loss and latency in the 
network already portends a problem. Overflowing queues in forwarding elements 
or mismatched hardware capabilities imply that traffic has not been optimally 
managed.
Ideally, a network should always be "zero-queue", i.e., latency 
will merely be induced by propagation, and not queuing delay.

TCP's algorithm is a best-effort solution to globally maximize utilization 
under the fairness principle. While this design has not fundamentally changed 
over the last thirty years, new research advancements raise the question if TCP 
is still the most reliable option.~\cite{pcc,bbr,perc}

@@

In parallel, admission-control-based approaches have gained traction as a 
viable alternative to the conventional burst-and-backoff mechanism of 
TCP.~\cite{expresspass, fastpass, perc}
While idea of admission control and service guarantees in networks is not 
new~\cite{access_limit, access_limit2}, these designs traditionally 
aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
and untrusted environments such as the internet. In a data center, these 
conditions do not apply. End-hosts are generally considered reliable and 
restricted in behaviour, which allows for great simplification of enforcement 
and prioritization policies. Admission-based systems can give administrators 
the advantage of tight control over the data center, mitigating the effects of 
queue-buildup and bursts.

Although it is unclear how predictable data center traffic truly is, various 
works have shown that repeated patterns may 
exist.\cite{msr_dc,fb_dc,traffic,microte}
If data center traffic exhibits predictable patterns, it may be 
possible to devise a proactive mechanism which forecasts the traffic matrix of 
next iterations, and adjusts traffic accordingly.

### 2017-12-11 Documents/Report/related_work.tex

\subsection{End-to-End TCP optimization}
Many different TCP optimizations tailored to various environments and use-cases 
exist. In general, these extensions share TCP's fundamental requirements of 
scalability, decentralization, and fairness. Optimizations are generally 
static, only on the end-host, and implemented "offline".~\cite{throwdown} This 
family of modifications includes tweaks to the congestion window size and 
growth, signal detection, and optimization goal(e.g., bandwidth, latency, or 
convergence).~\cite{tcp_family}

In this project, we compare Iroko to two particular TCP protocols of interest.  
Data Center TCP~\cite{dctcp}, a congestion control algorithm specialized to 
reduce bufferbloat in data center networks. DCTCP leverages Explicit Congestion 
Notifications (ECNs) as early warning signal to identify queue buildup and 
adjusts its sending rate to the amount of ECN-marked packets on the receiving 
side. As part of the static TCP congestion control family, DCTCP is fully 
decentralized, does not limit host sending rates in advance and relies on a 
static congestion control model.

The second algorithm we compare to is TCP CUBIC~\cite{cubic}, the default 
congestion control protocol of the Linux kernel. CUBIC uses a cubic window size 
function and serves as the baseline for Iroko.

@@

MicroTE~\cite{microte} is another central scheduler working under the 
assumption that a subset of short-term traffic is predictable. The controller 
isolates flows which it determined to be predictable and install a custom 
optimal routing entry. The remaining unpredictable traffic is routed according 
to ECMP striping.

@@

\subsection{Network Admission-Control}
Recent contributions have shown that purely admission-based approaches 
may work just as well as, if not better than, the conventional TCP fairness 
principle. A large focus of these systems is to eliminate queuing and fast flow 
convergence coupled with minimal round trip latency.
FastPass~\cite{fastpass} employs an arbiter mechanism, in which hosts request 
access to the network resources. For any new flow, the FastPass shim layer will 
send a "admission request" to a central controller, which chooses an optimal 
allocated time slot and path. Once an applications has been assigned this path, 
it is guaranteed to experience zero queuing latency and jitter during 
transmission. This "zero-loss" network provides application with a stable and 
reliable transmission experience. However, FastPass still works reactively and 
only operates in incoming requests. Queueing is thus pushed to end-hosts and 
the arbiter itself.

@@

\subsection{Predicting and Learning Network Traffic}
A new direction of networking research is study on the predictability and 
trainability of network traffic.~\cite{throwdown, learning_tcp}
A common assumption is that data center traffic is largely unpredictable and 
untrainable.~\cite{fb_dc}
However, Winstein and Balakrishnan,~2013\cite{remy} demonstrated that it is 
possible to learn near-optimal TCP congestion control parameters for various 
network topologies. The system, Remy, represents an offline learning algorithm 
inferring optimal TCP parameters based on pre-supplied topology and 
traffic-model.
While Remy only performed offline training, PCC~\cite{pcc,throwdown} aims to 
develop a dynamic and actively improving congestion control algorithm. PCC 
makes no prior assumption about the network and continuously reiterates over 
its host-local optimization function to find optimal parameters.
Furthermore, Benson et al.,~2014\cite{microte} analyzed data center traffic and 
inferred that around 30\% of short-lived traffic is predictable over a period 
of 1-2 seconds.

In this work, we assume that traffic may be trainable to a degree that is 
useful. We conjecture on the basis of the aforementioned work that it is 
possible to identify recurring patterns in a subset of traffic and adjust flow 
and congestion control pre-emptively.

>>>>

\subsection{End-to-End TCP optimization}
Many different TCP optimizations tailored to various environments and use-cases 
exist. In general, these extensions share TCP's fundamental requirements of 
scalability, decentralization, and fairness. Optimizations are generally only 
implemented on the end-host and the model itself is static and 
"offline".~\cite{throwdown} This family of modifications includes tweaks to the 
congestion window size and growth, signal detection, and optimization goal 
(e.g., bandwidth, latency, or convergence).~\cite{tcp_family}

In this project, we compare Iroko to two particular TCP protocols of interest.  
The first, Data Center TCP (DCTCP)~\cite{dctcp}, is a congestion control 
algorithm developed to reduce bufferbloat in data center networks. DCTCP 
leverages Explicit Congestion Notifications (ECNs) as early warning signal to 
identify queue buildup and adjusts its sending rate to the amount of ECN-marked 
packets on the receiving side. As part of the static TCP congestion control 
family, DCTCP is fully decentralized, does not limit host sending rates in 
advance and relies on a static congestion control model.

The second algorithm we compare to is TCP CUBIC~\cite{cubic}, the default 
congestion control protocol of the Linux kernel. CUBIC uses a cubic function   
to increase window size and serves as the baseline for Iroko. It is intended to 
work well over high bandwidth, high latency networks.

@@

MicroTE~\cite{microte} is a central scheduler working under the 
assumption that a subset of short-term traffic is predictable. The controller 
isolates flows which it determined to be predictable and install a custom 
optimal routing entry. The remaining unpredictable traffic is routed according 
to ECMP striping.

@@

\subsection{Network Admission-Control}
Recent contributions have shown that purely admission-based approaches 
may work as well as, if not better than, conventional unbounded TCP. A large 
focus of these systems is to eliminate queuing and fast flow 
convergence coupled with minimal round trip latency.

FastPass~\cite{fastpass} employs an arbiter mechanism, in which hosts request 
access to the network resources. For any new flow, the FastPass shim layer will 
send a "admission request" to a central controller, which chooses an optimal 
allocated time slot and path. Once an applications has been assigned this path, 
it is guaranteed to experience zero queuing latency and jitter during 
transmission. This "zero-loss" network provides application with a stable and 
reliable transmission experience. However, FastPass still works reactively and 
only operates in incoming requests. Queueing is thus pushed to end-hosts and 
the arbiter itself.

@@

\subsection{Predicting and Learning Network Traffic}
A new direction of networking research is study on the predictability and 
trainability of network traffic.~\cite{throwdown, learning_tcp}
A common assumption is that data center traffic is largely unpredictable and 
untrainable.~\cite{fb_dc}
However, Winstein and Balakrishnan,~2013\cite{remy} demonstrated that it is 
possible to learn near-optimal TCP congestion control parameters for various 
network topologies. The system, Remy, represents an offline learning algorithm 
inferring optimal TCP parameters based on pre-supplied topology and 
traffic-model.
While Remy only performed static offline training, PCC~\cite{pcc,throwdown} 
aims to develop a dynamic and actively improving congestion control algorithm. 
PCC makes no prior assumption about the network and continuously reiterates over
its host-local optimization function to find optimal parameters.

As part of the contribution of MicroTE, Benson et 
al.,~2014~\cite{microte} analyzed data center traffic and 
inferred that around 30\% of short-lived traffic is predictable over a period 
of 1-2 seconds.

In this work, we assume that traffic may be trainable to a degree that is 
useful. On the basis of the aforementioned work we conjecture that it may be 
possible to identify recurring patterns in a subset of traffic and adjust flow 
and congestion control pre-emptively.
